---
title: "Assignment 7"
output:
  word_document: default
  html_notebook: default
---

```{r}
library(rvest)
library(httr)
library(stringr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggpubr)
```

Question 1.
1. Scrape the data from the webpage and extract the following fields: Country, < $1.90, < $3.20, < $5.50, Year and Continent. Prepare the data for analysis and ensure that the columns have meaningful names.

```{r}
url = 'https://en.m.wikipedia.org/wiki/List_of_countries_by_percentage_of_population_living_in_poverty'
webpage = read_html(url)
wiki_table = webpage %>%
  html_nodes("table") %>%
  .[[2]] %>%
  html_table(header = TRUE, fill = TRUE)
```
The first chunk of the code is reading the table from the link mentioned.
In the chunk below we are changing the names of the columns to a more manageable format, as names before has space and special character which will be difficult to call in R.
```{r}
colnames(wiki_table) <- c('Country', 'less_than_1.90', 'less_than_3.20', 'less_than_5.50', 'Year', 'Continent')
str(wiki_table)
```
Through str(), three columns which should be integer in datatype are as character, and also it has % sign which will not allow us to perform operation on the data. So below is the code chunks that solves these problem.
```{r}
wiki_table2 <- wiki_table %>%
  separate(less_than_3.20, into = c("less_than_3.20"), sep = "%") %>%
  separate(less_than_1.90, into = c("less_than_1.90"), sep = "%") %>%
  separate(less_than_5.50, into = c("less_than_5.50"), sep = "%")
```
```{r}
wiki_table2$less_than_1.90 <- as.integer(wiki_table2$less_than_1.90)
wiki_table2$less_than_3.20 <- as.integer(wiki_table2$less_than_3.20)
wiki_table2$less_than_5.50 <- as.integer(wiki_table2$less_than_5.50)
str(wiki_table2)
```
Made a new dataframe named wiki_table2 so that there is no overlap of data while changing datatye and removing % sign and get a clean data.

2. Calculate the mean and the standard deviation of the percent of the population living under $5.50 per day for each continent. Perform a comparative analysis (i.e. explanation) of the data from each continent.

```{r}
wiki_table2 %>%
  group_by(Continent) %>%
  mutate(mean_percentage = mean(less_than_5.50), sd_percentage = sd(less_than_5.50)) %>%
  distinct(Continent, mean_percentage, sd_percentage)
```
Simple dplyr functions were applied to to extract mean and standard deviation of all continents.
Out of all continents, Europe has the least mean percentage of 4.2% population recieving daily wage below $5.50,and it also has the least standard deviation (8.8) that means the data is concentrated towards the mean value.

3. What are the 10 countries with the highest percentage of the population having an income of less than $5.50 per day? Using a suitable chart, display the country name, the percentage and color- code by the Continent. Summarize your findings.

```{r}
wiki_table2 %>%
  select(Country, less_than_5.50, Continent) %>%
  arrange(desc(less_than_5.50)) %>%
  slice(1:10) %>%
  ggplot(aes(x = Country, y = less_than_5.50, color = Continent)) + geom_point() + theme(axis.text.x = element_text(angle = 90) )
```
The 10 countries, Burundi, Central African Republic, Congo, Liberia, Madagascar, Malawi, Mozambique, Somalia, South Sudan, Uzbekistan are the top countries with highest percentage of population having income less than $5.5.
According to the graph, all countries are from Africa, except Uzbekistan which is an Asian country.

4. Explore the countries with the lowest percentage of the population having an income of less than $5.50 per day. What are the 5 countries with the lowest percentage, and how does the results compare to the other income groups (i.e. $1.90 and $3.20)

```{r}
low_5.5 <- wiki_table2 %>%
  select(Country, Year, less_than_1.90, less_than_3.20, less_than_5.50) %>%
  arrange(less_than_5.50) %>%
  slice(1:5) %>%
  arrange(Year)
low_1.9 <- wiki_table2 %>%
  select(Country, Year, less_than_1.90, less_than_3.20, less_than_5.50) %>%
  arrange(less_than_1.90) %>%
  slice(1:5) %>%
  arrange(Year)
low_3.2 <- wiki_table2 %>%
  select(Country, Year, less_than_1.90, less_than_3.20, less_than_5.50) %>%
  arrange(less_than_3.20) %>%
  slice(1:5) %>%
  arrange(Year)

plot_1.9 <- ggplot(data = low_1.9, mapping = aes(x = Year)) + geom_line(mapping = aes(y = less_than_1.90), color = "blue") + geom_point(mapping = aes(y = less_than_1.90, color = Country)) +
  geom_line(mapping = aes(y = less_than_3.20), color = "red") +
  geom_point(mapping = aes(y = less_than_3.20, color = Country)) +
  geom_line(mapping = aes(y = less_than_5.50), color = "green") +
  geom_point(mapping = aes(y = less_than_5.50, color = Country)) + 
  labs(y = "Percentage of Population", title = "Data of Top 5 Countries with < $1.90 Income Per Day")

plot_3.2 <- ggplot(data = low_3.2, mapping = aes(x = Year)) + geom_line(mapping = aes(y = less_than_1.90), color = "blue") + geom_point(mapping = aes(y = less_than_1.90, color = Country)) +
  geom_line(mapping = aes(y = less_than_3.20), color = "red") +
  geom_point(mapping = aes(y = less_than_3.20, color = Country)) +
  geom_line(mapping = aes(y = less_than_5.50), color = "green") +
  geom_point(mapping = aes(y = less_than_5.50, color = Country)) + 
  labs(y = "Percentage of Population", title = "Data of Top 5 Countries with < $3.20 Income Per Day")

plot_5.5 <- ggplot(data = low_5.5, mapping = aes(x = Year)) + geom_line(mapping = aes(y = less_than_1.90), color = "blue") + geom_point(mapping = aes(y = less_than_1.90, color = Country)) +
  geom_line(mapping = aes(y = less_than_3.20), color = "red") +
  geom_point(mapping = aes(y = less_than_3.20, color = Country)) +
  geom_line(mapping = aes(y = less_than_5.50), color = "green") +
  geom_point(mapping = aes(y = less_than_5.50, color = Country)) + 
  labs(y = "Percentage of Population", title = "Data of Top 5 Countries with < $5.50 Income Per Day")

ggarrange(plot_1.9, plot_3.2, plot_5.5)
```
Countries with the lowest percentage of the population having an income of less than $5.50 per day are Australia, Austria, Belarus, Belgium, and Canada. For this filtered data, percentage of the population having an income of less than $1.9 and $3.2 per day is also zero.
However, this is not the case for other two cases as can be seen in graph.

5. Extract the data for any two continents of your choice. Fore each continent, visualize the percent of the population living on less than $1.90, $3.20 and $5.50 using box plots. Compare and contrast the results, while ensuring that you discuss the distribution, skew and any outliers that are evident.

```{r}
box_1.90 <- wiki_table2 %>%
  filter(Continent=='Asia'| Continent=='Oceania') %>%
  ggplot(aes(x = Continent, y = less_than_1.90)) + geom_boxplot()
box_3.20 <- wiki_table2 %>%
  filter(Continent=='Asia' | Continent=='Oceania') %>%
  ggplot(aes(x = Continent, y = less_than_3.20)) + geom_boxplot()
box_5.50 <- wiki_table2 %>%
  filter(Continent=='Asia' | Continent=='Oceania') %>%
  ggplot(aes(x = Continent, y = less_than_5.50)) + geom_boxplot()
```
```{r}
ggarrange(box_1.90, box_3.20, box_5.50)
```
Box plot of less than 1.9 shows Asia to have minimum distribution and and all data concentrated towards mean, (~0), but it has outliers going till 60%. In the third plot for less than 5.5 Asia has maximum distribution and has no outliers.
